---
title: 'STAT 224 Autumn 2022 Homework 2'
author: "Matthew Zhao"
fontsize: 12pt
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes: \usepackage{pdfpages}
geometry: margin=0.75in
fig_crop: no
---

```{r echo=F, include = FALSE}
library(knitr)
knitr::opts_chunk$set(fig.width=4, fig.height=3, message=F, warning=F, dpi=36)
options(width=70, digits=4, scipen=8)
```

\newcommand{\ansbegin}{\smallbreak\vspace{-8pt}\hrulefill \smallbreak\noindent {\sl Answer\/}:\enspace\ignorespaces}
\newcommand{\ansend}{\smallbreak\hrulefill \smallbreak\vspace{-8pt}}

<!-- **Due midnight on Friday, October 15, 2021** on Gradescope. -->

# Question 1


```{r}
fevdata = read.table("http://www.stat.uchicago.edu/~yibi/s224/data/fevdata.txt", h = TRUE)
fevdata$sex = factor(fevdata$sex, labels=c("Female","Male"))
fevdata$smoke = factor(fevdata$smoke, labels=c("Nonsmoker","Smoker"))
```

## Q1a --- 6 points

```{r}
lmm.nosmoke = lm(fev ~ age, data=subset(fevdata, sex == "Male" & smoke == "Nonsmoker"))
summary(lmm.nosmoke)
```

$\hat{\beta}_{0}^{mn} = -0.0576$, $s.e.(\hat{\beta}_{0}^{mn}) = 0.1147$

$\hat{\beta}_{1}^{mn} = 0.2882$, $s.e.(\hat{\beta}_{1}^{mn}) = 0.0114$

$\hat{\sigma}^{mn} = 0.556, n=654$

## Q1b --- 8 points

$t=\frac{\hat{\beta}_j - \beta_{j}^0}{s.e.({\hat{\beta}_j})}$

### i)

```{r}
t = (-0.0576 - 0)/0.1147
n=654
p=1
pval = pt(t,df=n-p-1,lower.tail = F)
print(paste0("t-stat: ",signif(t,digits = 5)))
print(paste0("df: ",n-p-1))
print(paste0("P-value: ",signif(pval,digits = 5)))
```

### ii)

```{r}
t = (-0.0576 - 0.1)/0.1147
n=654
p=1
pval = 2*pt(abs(t),df=n-p-1,lower.tail = F)
print(paste0("t-stat: ",signif(t,digits = 5)))
print(paste0("df: ",n-p-1))
print(paste0("P-value: ",signif(pval,digits = 5)))
```

### iii)

```{r}
t = (0.2882 - 0.1)/0.0114
n=654
p=1
pval = pt(t,df=n-p-1)
print(paste0("t-stat: ",signif(t,digits = 5)))
print(paste0("df: ",n-p-1))
print(paste0("P-value: ",signif(pval,digits = 5)))
```

### iv)

```{r}
t = (0.2882 - 0.3)/0.0114
n=654
p=1
pval = pt(t,df=n-p-1,lower.tail = F)
print(paste0("t-stat: ",signif(t,digits = 5)))
print(paste0("df: ",n-p-1))
print(paste0("P-value: ",signif(pval,digits = 5)))
```

## Q1c --- 4 points

$\hat{\beta}_j \pm t_{(n-p-1,\frac{\alpha}{2})} * s.e.(\hat{\beta}_j)$

```{r}
beta_hat = 0.2882 
se = 0.0114
n=654
p=1
alpha = 0.1
t = qt(alpha/2,df=n-p-1,lower.tail = F)
print(paste0("confidence interval: ",'(',
             signif(beta_hat-t*se,digits=5),',', 
             signif(beta_hat+t*se,digits=5),')'))
```

We are 90% confident that $\beta_{1}^{mn}$ is between 0.26942 and 0.30698.

## Q1d --- 5 points

```{r, message=FALSE}
aggregate(age ~ sex + smoke, data=fevdata, mean)
aggregate(age ~ sex + smoke, data=fevdata, sd)
```


## Q1e --- 4 points



## Q1f --- 5 points



## Q1g --- 6 points



# Question 2

```{r}
NLSY = read.table("http://www.stat.uchicago.edu/~yibi/s224/data/NLSY.txt", header=T)
NLSYm = subset(NLSY, Gender == "male")
```

## Q2a --- 6 points

```{r}
lm1 = lm(log(Income2005) ~ AFQT, data=NLSYm)
lm2 = lm(log(Income2005) ~ AFQT + Edu2006, data=NLSYm)
```


## Q2b --- 5 points



## Q2c --- 2 points



