---
title: 'STAT 224 Autumn 2022 HW3'
author: "Matthew Zhao"
fontsize: 12pt
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes: \usepackage{pdfpages}
geometry: margin=0.75in
fig_crop: no
---

```{r echo=F, include = FALSE}
library(knitr)
knitr::opts_chunk$set(fig.width=4, fig.height=3, message=F, warning=F)
options(width=70, digits=4, scipen=8)
```


```{r}
food = read.table("http://www.stat.uchicago.edu/~yibi/s224/data/food.txt", h = T)
ee = read.table("http://www.stat.uchicago.edu/~yibi/s224/data/EE.txt", h = T)
te = read.table("http://www.stat.uchicago.edu/~yibi/s224/data/TE.txt", h = T)
```

## Q1a --- 3 Points

```{r}
lmte = lm(log(V) ~ log(K) + log(L), data = te)
print(lmte$coefficients[2:3])
confint(lmte,'log(L)',level = 0.90)
```

Least Squares estimates for $\beta_K$ and $\beta_L$ are 0.5057 and 0.8455 respectively, with the 90% CI for $\beta_L$: (0.09411,1.597). This means that we are 90% confident that a 1% increase in labor input would lead to an increase in output of 0.09411 to 1.597%. 

## Q1b --- 5 points


```
> summary(lmte)
Residual standard error: 0.07495 on 12 degrees of freedom
Multiple R-squared:  0.8184,	Adjusted R-squared:  0.7881 
$F$-statistic: 27.03 on 2 and 12 DF,  p-value: 3.591e-05
```

$SSE = 0.07495$

$R^2 = 1 - \frac{SSE}{SST}, SST = SSR + SSE \Rightarrow R^2 = 1-\frac{SSE}{SSR + SSE} \Rightarrow SSR = SSE \times \frac{R^2}{1-R^2} \Rightarrow SSR = 0.07495 \times \frac{0.8184}{1-0.8184} = 0.3378$

Since the F-statistic is drawn from F distribution $F_{2,12}$, dfR = 2 and dfE = 12 so 
$MSR = \frac{SSR}{dfR} = \frac{0.07495}{2} = 0.03748$ and 
$MSE = \frac{SSe}{dfE} = \frac{0.3378}{12} = 0.02815$

The F-statistic = 27.03

```{r}
pf(27.03,2,12,lower.tail = F)
```

Source | df | SS | MS | F | P-value 
:--------|--:|:--------:|:-------:|:------:|------:
Regression | 2 | SSR = 0.3378 | MSR = 0.03748 | F = 27.03 | 0.00003593 
Error | 12 | SSE = 0.07495 | MSE = 0.02815 |  | 


## Q1c --- 5 points

Full Model: $\log{V} = \log{\alpha} + \beta_K \log{K} + \beta_{L} \log{L} + \epsilon$

Reduced Model ($\beta_K = \beta_L$): $\log{V} = \log{\alpha} + \beta_K (\log{K} + \log{L}) + \epsilon$

```{r}
full_lmte = lm(log(V) ~ log(K) + log(L), data = te)
reduced_lmte = lm(log(V) ~ I(log(K) + log(L)), data = te)
anova(reduced_lmte,full_lmte)
```

$F = \frac{(SSE_{reduced} - SSE_{full})/(dfE_{reduced} - dfE_{full})}{MSE_{full}} = \frac{(0.0682-0.0674)/(13-12)}{0.0674/12} = 0.1424$

Degrees of freedom for F-stat is $dfE_r - dfE_f$ and $dfE_f$ so 13-12 = 1 and 12. 

```{r}
pf(0.1424,1,12,lower.tail = F)
```

Since the F-test is not statistically significant, we say that we fail to reject the null that $\beta_L = \beta_K$. 

## Q1d --- 6 Points

```{r}
full_lm = lm(log(V) ~ log(K) + log(L), data = te)
reduced_lm = lm(log(V) ~ I(log(K)-log(L)), offset=log(L), data = te)
anova(reduced_lm,full_lm)
```

For Transportation Equipment (TE): F-stat 2.87 with $dfE_r - dfE_f = 1$ and $dfE_f = 12$, p-value 0.12

```{r}
full_lm = lm(log(V) ~ log(K) + log(L), data = ee)
reduced_lm = lm(log(V) ~ I(log(K)-log(L)), offset=log(L), data = ee)
anova(reduced_lm,full_lm)
```

For Equipment and Supplies (EE): F-stat 0.41 with $dfE_r - dfE_f = 1$ and $dfE_f = 12$, p-value 0.53.

At the 0.05 significance level, we fail to reject that $\beta_L + \beta_K = 1$

## Q1e --- 2 Points


```{r}
model = lm(log(V) ~ log(K) + log(L) + YEAR, data = ee)
summary(model)
```

Since $H_0: \log{\rho} =0$ and $H_a: \log{\rho} >0$, we use a 1-sided t-test.

$t = \frac{\hat{\beta}_j - \beta_{j}^0}{se(\hat{\beta}_j)} = \frac{0.02497-0}{0.00346} = 7.217$

```{r}
pt(7.217,nrow(ee)-3-1,lower.tail = F)
```

Since the p-value is less than 0.05, we reject the null hypothesis, meaning that there was technological progress from 1972 to 1986 for the EE sector.

## Q1f --- 5 points

```{r}
full_food = lm(log(V) ~ log(K) + log(L) + YEAR, data = food)
reduced_food = lm(log(V) ~ YEAR, data = food)
anova(reduced_food, full_food)
```

$F = \frac{(SSE_{reduced}-SSE_{full})/(dfE_{reduced}-dfE_{full})}{MSE_{full}} = \frac{(0.0447-0.0413)/(13-11)}{0.0413/11} = 0.4527$

$dfE_{reduced}-dfE_{full} = 13-11 = 2$, $dfE_{full} = 11$.

```{r}
pf(0.4527,2,11,lower.tail = F)
```

Since the p-value is higher than 0.05, we fail to reject the null and conclude that $\beta_K = \beta_L = 0$


# Question 2


## Q2a --- 3 points



## Q2b --- 2 points



## Q2c --- 2 points



## Q2d --- 2 points



## Q2e --- 3 points



## Q2f --- 4 points



## Q2g --- 2 points



## Q2h --- 4 points



## Q2i --- 2 points

