---
title: 'STAT 224 Autumn 2022 HW3'
author: "Matthew Zhao"
fontsize: 12pt
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes: \usepackage{pdfpages}
geometry: margin=0.75in
fig_crop: no
---

```{r echo=F, include = FALSE}
library(knitr)
knitr::opts_chunk$set(fig.width=4, fig.height=3, message=F, warning=F)
options(width=70, digits=4, scipen=8)
```


```{r}
food = read.table("http://www.stat.uchicago.edu/~yibi/s224/data/food.txt", h = T)
ee = read.table("http://www.stat.uchicago.edu/~yibi/s224/data/EE.txt", h = T)
te = read.table("http://www.stat.uchicago.edu/~yibi/s224/data/TE.txt", h = T)
```

## Q1a --- 3 Points

```{r}
lmte = lm(log(V) ~ log(K) + log(L), data = te)
print(lmte$coefficients[2:3])
confint(lmte,'log(L)',level = 0.90)
```

Least Squares estimates for $\beta_K$ and $\beta_L$ are 0.5057 and 0.8455 respectively, with the 90% CI for $\beta_L$: (0.09411,1.597). This means that we are 90% confident that a 1% increase in labor input would lead to an increase in output of 0.09411 to 1.597%. 

## Q1b --- 5 points


```
> summary(lmte)
Residual standard error: 0.07495 on 12 degrees of freedom
Multiple R-squared:  0.8184,	Adjusted R-squared:  0.7881 
$F$-statistic: 27.03 on 2 and 12 DF,  p-value: 3.591e-05
```

$SSE = 0.07495$

$R^2 = 1 - \frac{SSE}{SST}, SST = SSR + SSE \Rightarrow R^2 = 1-\frac{SSE}{SSR + SSE} \Rightarrow SSR = SSE \times \frac{R^2}{1-R^2} \Rightarrow SSR = 0.07495 \times \frac{0.8184}{1-0.8184} = 0.3378$

Since the F-statistic is drawn from F distribution $F_{2,12}$, dfR = 2 and dfE = 12 so 
$MSR = \frac{SSR}{dfR} = \frac{0.07495}{2} = 0.03748$ and 
$MSE = \frac{SSe}{dfE} = \frac{0.3378}{12} = 0.02815$

The F-statistic = 27.03

```{r}
pf(27.03,2,12,lower.tail = F)
```

Source | df | SS | MS | F | P-value 
:--------|--:|:--------:|:-------:|:------:|------:
Regression | 2 | SSR = 0.3378 | MSR = 0.03748 | F = 27.03 | 0.00003593 
Error | 12 | SSE = 0.07495 | MSE = 0.02815 |  | 


## Q1c --- 5 points



## Q1d --- 6 Points



## Q1e --- 2 Points



## Q1f --- 5 points



# Question 2


## Q2a --- 3 points



## Q2b --- 2 points



## Q2c --- 2 points



## Q2d --- 2 points



## Q2e --- 3 points



## Q2f --- 4 points



## Q2g --- 2 points



## Q2h --- 4 points



## Q2i --- 2 points

