---
title: 'STAT 224 Autumn 2022 HW6'
author: "Matthew Zhao"
fontsize: 12pt
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes: \usepackage{pdfpages}
geometry: margin=0.5in
fig_crop: no
---

```{r echo=F, include = FALSE}
library(knitr)
library(ggplot2)
knitr::opts_chunk$set(fig.width=4, fig.height=3, message=F, warning=F, collapse=TRUE)
options(width=70, digits=4, scipen=8)
options(show.signif.stars = FALSE) # Show no stars for significance tests
```

# Question 1

http://www.stat.uchicago.edu/~yibi/s224/data/brake.txt

```{r load-brake}
brake = read.table("http://www.stat.uchicago.edu/~yibi/s224/data/brake.txt", header=T)
```

## Q1a --- 3 points

```{r}
ols1 = lm(distance ~ speed + I(speed^2), data=brake)
ggplot(data=brake,aes(x=speed,y=distance)) + 
  geom_point() + labs(x='speed of car (mph)',y='distance to stop (ft)')
ggplot(data=brake,aes(x=ols1$fitted.values,y=ols1$residuals)) + 
  geom_point() + labs(x='fitted values',y='raw residuals')
```

Yes, there is clearly nonconstant variance. 

## Q1b --- 3 points


```{r}
ols2 = lm(sqrt(distance) ~ speed + I(speed^2), data=brake)
ols3 = lm(log(distance) ~ speed + I(speed^2), data=brake)
```
```{r}
ggplot(data=brake,aes(x=ols2$fitted.values,y=ols2$residuals)) + 
  geom_point() + labs(x='fitted values',y='raw residuals')
ggplot(data=brake,aes(x=ols3$fitted.values,y=ols3$residuals)) + 
  geom_point() + labs(x='fitted values',y='raw residuals')
```

Square-root appears to be the most appropriate transformation since this transformation results in near constant variance based on the residual plots. 

## Q1c --- 2 points


```{r}
library(MASS)
boxcox(ols1)
abline(v=1/2, col="red") 
```

Since $\frac{1}{2}$ falls within the 95% CI, squart root is the most appropriate transformation. 

## Q1d --- 2 points

```{r, fig.width=2.5, fig.height=2.5}
predCI = predict(ols1, data.frame(speed = 4:40), interval="prediction")
predCI = data.frame(x=4:40,predCI)
ggplot() +
  geom_ribbon(data=predCI, aes(x=x, ymin=lwr, ymax = upr), alpha=0.25) +
  geom_line(data=predCI, aes(x=x,y = fit),col="blue",lwd=1) +
  geom_point(data=brake, aes(x=speed, y=distance)) +
  labs(x="Auto Speed (mph)", y="Distance Until Stop (ft.)")
```

```{r, fig.width=2.5, fig.height=2.5}
predCI = predict(ols2, data.frame(speed = 4:40), interval="prediction")^2
predCI = data.frame(x=4:40,predCI)
ggplot() +
  geom_ribbon(data=predCI, aes(x=x, ymin=lwr, ymax = upr), alpha=0.25) +
  geom_line(data=predCI, aes(x=x,y = fit),col="blue",lwd=1) +
  geom_point(data=brake, aes(x=speed, y=distance)) +
  labs(x="Auto Speed (mph)", y="Distance Until Stop (ft.)")
```

```{r, fig.width=2.5, fig.height=2.5}
predCI = exp(predict(ols3, data.frame(speed = 4:40), interval="prediction"))
predCI = data.frame(x=4:40,predCI)
ggplot() +
  geom_ribbon(data=predCI, aes(x=x, ymin=lwr, ymax = upr), alpha=0.25) +
  geom_line(data=predCI, aes(x=x,y = fit),col="blue",lwd=1) +
  geom_point(data=brake, aes(x=speed, y=distance)) +
  labs(x="Auto Speed (mph)", y="Distance Until Stop (ft.)")
```

The prediction band for ols2 best matches the pattern of the data. 


## Q1e --- 4 points

```{r}
summary(ols2)
```

It is not significant. 

```{r}
ols2_nosquare = lm(sqrt(distance) ~ speed, data=brake)
boxcox(ols2_nosquare)
library(car)
crPlots(ols2_nosquare,'speed')
```

The 95% CI for box-cox contains 1 indicating that no further transformation is needed. Additionally we see no signs of nonlinearity via the residual plus component plot for our only covariate speed. 

## Q1f --- 3 points

```{r}
predCI = predict(ols2, data.frame(speed = seq(10,30,10)), interval="prediction")^2
data.frame(speed=seq(10,30,10),predCI)
```

# Question 2


## Q2a --- 4 points


```{r}
wls1 = lm(distance ~ speed + I(speed^2), data=brake, weight=1/speed)
wls2 = lm(distance ~ speed + I(speed^2), data=brake, weight=1/speed^2)
```
```{r}
ggplot(data=brake,aes(x=speed,y=rstandard(wls1))) + 
  geom_point() + 
  labs(x='speed',y='raw residuals')
ggplot(data=brake,aes(x=speed,y=rstandard(wls2))) + 
  geom_point() + 
  labs(x='speed',y='raw residuals')
ggplot(data=brake,aes(x=speed,y=rstudent(wls2))) + 
  geom_point() + 
  labs(x='speed',y='raw residuals')

```

We should use the standardized (internally standardized) residuals for model 2. 

## Q2b --- 2 points

```{r}
predCI = predict(wls2, data.frame(speed = seq(10,30,10)), interval="prediction")
data.frame(speed=seq(10,30,10),predCI)
```

# Question 3


http://www.stat.uchicago.edu/~yibi/s224/data/fatherson.txt

```{r}
fatherson = read.table("http://www.stat.uchicago.edu/~yibi/s224/data/fatherson.txt", h=T)
```


## Q3a --- 3 points


```{r}
lm1 = lm(Aver.Height.Son ~ Height.Father, data=fatherson)
```

```{r}
ggplot(data=fatherson,aes(x=lm1$fitted.values,y=lm1$residuals)) + 
  geom_point() + 
  labs(x='fitted values',y='raw residuals')
crPlots(lm1,'Height.Father')
boxcox(lm1)
```

OLS is not appropriate since the residual plot reveals that nonconstant variance is violated. Additionally, boxcox shows that there are no ideal transformations of the response to solve this issue since many $\lambda$s fall within the 95% CI.

## Q3b --- 6 points

```{r}
wls3 = lm(Aver.Height.Son ~ Height.Father, data=fatherson,
          weight=1/(Height.Father^(1/2)))
summary(wls3)
ggplot(data = fatherson,aes(x=Height.Father,y=rstandard(wls3))) + 
  geom_point() + labs(x='Father Height',y='standardized residuals') + 
  geom_hline(yintercept = 0)
```

## Q3c --- 2 points



# Question 4


http://www.stat.uchicago.edu/~yibi/s224/data/P229-30.txt

```{r}
stock = read.table("http://www.stat.uchicago.edu/~yibi/s224/data/P229-30.txt", header=T)
```


## Q4a --- 8 points

```{r}
stock1 = lm(DJIA ~ Day, data=stock)
```

```{r}
ggplot(data=stock,aes(x=Day,y=DJIA)) + 
  geom_point() + labs(x='Day',y='DJIA Index')
ggplot(data=stock,aes(x=Day,y=stock1$residuals)) + 
  geom_line() + labs(x='Time (day)','Raw Residuals') + 
  geom_hline(yintercept = 0)
library(tseries)
runs.test(factor(stock1$residuals > 0))
durbinWatsonTest(stock1)
acf(stock1$residuals)
```

The model does exhibit time dependencies based on the number of runs and small p values for the runs test, durbin watson test, and autocorrelation plot.

## Q4b --- 8 points

```{r}
stock2 = lm(DJIA[2:262] ~ DJIA[1:261], data=stock)
```

```{r}
ggplot(mapping=aes(x=stock$Day[1:261],y=stock2$residuals)) + 
  geom_line() + labs(x='Time (day)','Raw Residuals') + 
  geom_hline(yintercept = 0)
library(tseries)
runs.test(factor(stock2$residuals > 0))
durbinWatsonTest(stock2)
acf(stock2$residuals)
```

There is no longer any evidence of autocorrelation in the residuals. 

