---
title: |
       | STAT 224 Intro to R Markdown
author: "Yandi Shen"
params:
  dotabs: no
output: 
  beamer_presentation:
    #theme: "metropolis"
    keep_tex: true
    highlight: "default"
    includes:
      in_header: "lec_style.tex"
    slide_level: 2
---

```{r setup, echo = FALSE, include = FALSE}
options(scipen=8, digits=4)
knitr::opts_chunk$set(
  size = 'small',
  comment=NA,
  collapse =TRUE
)

###########
# improving ggplot2 appearance on a projector screen
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(panel.grid.major = ggplot2::element_line(colour = "grey75"))
```

## 

Please download the files below from Canvas

* `LA0928_demo1.Rmd`
* `LA0928_demo2.Rmd`
* `LA0928_demo3.Rmd`

## Itemized List

\begin{flushright}
\fbox{\begin{minipage}{0.6\textwidth}
\begin{itemize}
\itemsep=0pt
\item Leave a blank line before the list
\item Indent using at least 4 spaces
\end{itemize}
\end{minipage}}
\end{flushright}

\vspace{-40pt}
R Markdown
```
* item 1
* item 2
    + sub-item 1
    + sub-item 2
* item 3
    + sub-item 1
```

**Knitted Output:**

* item 1
* item 2
    + sub-item 1
    + sub-item 2
* item 3
    + sub-item 1

## Ordered List

\begin{flushright}
\fbox{\begin{minipage}{0.6\textwidth}
\begin{itemize}
\itemsep=0pt
\item Leave a blank line before the list
\item Indent using at least 4 spaces
\end{itemize}
\end{minipage}}
\end{flushright}

\vspace{-40pt}
**R Markdown:**
```
1. item 1
2. item 2
    a. sub-item 1
    b. sub-item 2
3. item 3
    a. sub-item 1
```

Knitted Output

1. item 1
2. item 2
    a. sub-item 1
    b. sub-item 2
3. item 3
    a. sub-item 1

## Bold-face, Italic, Striking Through Text

**R Markdown:**

```
1. **This text is bold.** 
1. *This text is in italics*.
1. You can even ~~strikethrough text~~
```

**Knitted Output:** 

1. **This text is bold.** 
1. *This text is in italics*.
1. You can even ~~strikethrough text~~

## Header

```
# Header 1
## Header 2
### Header 3

ordinary text
```

## Table

**R Markdown:**

```
Table Header  | Second Header
------------- | -------------
Table Cell 1  | Cell 2
Cell 3        | Cell 4
```

**Knitted Output:**

Table Header  | Second Header
------------- | -------------
Table Cell 1  | Cell 2
Cell 3        | Cell 4

## 

You can center, right-, or left-justify  text in your tables as you like.

**R Markdown:**

```
Centered | Right-Justified | Left-Justified
:-------:|----------------:|:----------------
A        | 24              |  My Friend's Name
B        |  5              |  My Name
CC       | 167             |  Firstname Lastname
DDD      | 48              |  Another Name
```
**Knitted Output:**

Centered | Right-Justified | Left-Justified
:-------:|----------------:|:----------------
A        | 24              |  My Friend's Name
B        |  5              |  My Name
CC       | 167             |  Firstname Lastname
DDD      | 48              |  Another Name


## Inline Math Symbols & Expressions (\LaTeX)

R Markdown supports \LaTeX  math symbols & expressions.
Inline math symbols & expressions are placed between `$ ... $`.

**R markdown:**

```
You can write Greek letters: $\alpha$, $\beta$, $\gamma$, $\delta$, 
$\epsilon$, $\varepsilon$ ...

**Subscripts** are written as $a_{b}$ and **superscripts** as $a^{b}$.

Simple Linear Regression Model: $Y_{ij}=\beta_0+\beta_1 X +\varepsilon_{ij}$
```

**Knitted Output:**

You can write Greek letters: $\alpha$, $\beta$, $\gamma$, $\delta$, 
$\epsilon$, $\varepsilon$ ...

**Subscripts** are written as $a_b$ and **superscripts** as $a^b$.

Simple Linear Regression Model: $Y_{ij}=\beta_0+\beta_1 X +\varepsilon_{ij}$

## 
\small
Place **two dollar signs** around math expressions in **display mode** like `$$... $$`

**R Markdown**
```
**Fractions** in **inline** math mode: $\frac{a}{b}$ and 
in  **display** math mode: $$\frac{a}{b}$$

Logistic Regression Model:
$$
P(Y=1) = \frac{e^{\alpha+\beta x}}{1+e^{\alpha+\beta x}}
$$
```

**Knitted Output:**

**Fractions** in **inline** math mode: $\frac{a}{b}$ and 
in  **display** math mode: $$\frac{a}{b}$$

Logistic Regression Model:
$$
P(Y=1) = \frac{e^{\alpha+\beta x}}{1+e^{\alpha+\beta x}}
$$

## Log, Exp, Square-root, and Summation


**R Markdown:**
```
$\log(x)$, $\exp(x)$, $\sqrt{x}$, $\sum_{i=1}^{n}x_i$
```
**Knitted Output:**

$\log(x)$, $\exp(x)$, $\sqrt{x}$, $\sum_{i=1}^{n}x_i$


Be sure to place the black-slash `\`, or they would look like

**R Markdown:**
```
$log(x)$, $exp(x)$, $sqrt{x}$, $sum_{i=1}^{n}x_i$
```
**Knitted Output:**

$log(x)$, $exp(x)$, $sqrt{x}$, $sum_{i=1}^{n}x_i$


## A \LaTeX  Tutorial
\small

`https://www.overleaf.com/learn/latex/Learn_LaTeX_in_30_minutes`

You can just look at the section "Adding math to \LaTeX" and 
all the sub-sections underneath.

* Mathematical expressions
* Subscripts and superscripts
* Brackets and Parentheses
* Fractions and Binomials
* Aligning Equations
* Operators
* Spacing in math mode
* Integrals, sums and limits
* Display style in math mode
* List of Greek letters and math symbols
* Mathematical fonts

## R Code Chunks

An R code chunk begin with 3 **back-tick** and then `{r}` and end with another
3 **back-tick**

```{r, results='asis', echo=FALSE}
cat("`` ```{r}  ``")
```

\vspace{-12pt}
```
x = c(1,2,3,5)
mean(x)
```

\vspace{-12pt}
```{r, results='asis', echo=FALSE}
cat("`` ```  ``")
```

\begin{center}
\includegraphics[width=0.85\textwidth]{./figures0928/Keyboard_US_back_tick.jpg}
\end{center}

## 

Use `echo=TRUE` or `echo=FALSE` to control whether to display the R codes.

Use `eval=TRUE` or `eval=FALSE` to control whether to evaluate the code chunks.

## The Iris Data

```{r}
data(iris)
str(iris)
```

## R Plots

```{r}
library(ggplot2)
ggplot(iris, aes(x=Petal.Width, y=Petal.Length))+
  geom_point()
```

## How to Adjust Size of R Plots

Adjust you can resize your plots by adjusting `fig.width` 
and `fig.height` in the code chunk.

Use `out.width` to adjust the relative width of the plot
to the document

See `LA0920_demo3.Rmd`.

## Center, Left- or Right-Justify R Plots

Use `fig.align='center'` to center your plot

Use `fig.align='right'` or `fig.align='left'` to left- or right-justify your plot.

See `LA0920_demo3.Rmd`.

## Placing Two Plots Side-by-Side

Use `fig.show='hold'` and  `out.width` 

See `LA0920_demo3.Rmd`.